Tutorial: SQL-on-Hadoop Systems

### 2015-04-01 paper.tex
updated refs and fixed some issues

\numberofauthors{4}
% 1st author
\author{
\alignauthor
Daniel Abadi\\
\affaddr{Yale University}\\
\email{daniel.abadi@yale.edu}
\alignauthor 
\newcommand{\spmarcel}{\hspace{-8mm}}
\spmarcel Shivnath Babu\\
\spmarcel \affaddr{Duke University}\\
\spmarcel \affaddr{Durham, NC} \\
\email{\spmarcel shivnath@cs.duke.edu}
\alignauthor
Fatma {\"O}zcan\\
\affaddr{IBM Research}\\
\email{fozcan@us.ibm.com}
\alignauthor
Ippokratis Pandis\\
\affaddr{Cloudera}\\
\email{ippokratis@cloudera.com}
}
\maketitle
% ----------------
% --- Abstract ---
% ----------------
%\begin{abstract}
%\end{abstract}
\section{Introduction}
Enterprises are using Hadoop as a central data repository for all their data coming from various sources,  including operational systems, social media and the web, sensors and smart devices, as well as their applications. SQL processing over Hadoop data in particular gained significant traction over the couple of years, as many enterprise data management tools rely on SQL,
and many enterprise users are familiar and comfortable with it. As a result, the number of SQL-on-Hadoop systems have increased significantly.

@@

\begin{itemize}
\item {{\bf Native systems with big data run-times:} These are systems that utilize a big data run-time such as Tez, or Spark to execute SQL queries. This category includes systems such as Hive and Spark SQL.}
\item{ {\bf Native systems with MPP architectures:} These are systems that employ an MPP database architecture that use stand-alone long running processing to execute SQL queries over HDFS data. This category includes systems such as impala, tajo, presto, and Big SQL}
\item{{\bf Database hybrids:} These are systems that combine a relational SQL engine with Hadoop run-time. This category includes systems such as Hadapt, HAWQ, and Vortex.}
\item{ {\bf Database connectors:} This category includes the extensions provided by major database vendors to reach out to Hadoop and ingest HDFS data. }
\item{ {\bf noSQL/Hbase specific variants:} These are systems that provide a declarative SQL layer on top of noSQL systems, HBase in particular. The systems in this category are Phoenix, and Splice Machine. Note that these systems focus on OLTP-like workloads, whereas all other categories address OLAP-like analytics}
\end{itemize}

@@

Apache Drill is an open-source project which aims at providing SQL-like declarative processing over self-describing semi-structured data. Its focus is on analyzing data without emposing a fixed schema or creating tables in a catalog like Hive MetaStore. It runs queries over files and HBase tables, and discovers data when reading input data. For every fixed chunk of data, it discovers its schema, creates an in-memory columnar representation, and generates specific code for processing. As such, it can accommodate data chunks with varying schemas. Drill provides a columnar execution engine that delays materialization as late as possible.

>>>>

\numberofauthors{4}

% 1st author
\author{
\alignauthor
Daniel Abadi\\
\affaddr{Yale University}\\
\email{daniel.abadi@yale.edu}
\alignauthor 
\newcommand{\spshiv }{\hspace{-12mm}}
\spshiv Shivnath Babu\\
\spshiv \affaddr{Duke University}
\spshiv \email{shivnath@cs.duke.edu} 
\alignauthor
\newcommand{\spfatma}{\hspace{-15mm}}
\spfatma Fatma {\"O}zcan\\
\spfatma \affaddr{IBM Research}
\spfatma \email{fozcan@us.ibm.com}
\alignauthor
\newcommand{\spippo}{\hspace{-12mm}}
\spippo  Ippokratis Pandis\\
\spippo \affaddr{Cloudera}
\spippo \email{ippokratis@cloudera.com}
}
\maketitle
% ----------------
% --- Abstract ---
% ----------------
%\begin{abstract}
%\end{abstract}
\section{Introduction}
Enterprises are using Hadoop as a central data repository for all their data coming from various sources,  including operational systems, social media and the web, sensors and smart devices, as well as their applications. SQL processing over Hadoop data in particular gained significant traction over the couple of years, as many enterprise data management tools rely on SQL,
and many enterprise users are familiar and comfortable with it. As a result, the number of SQL-on-Hadoop systems have increased significantly.

@@

\begin{itemize}
\item {{\bf Native systems with big data run-times:} These are systems that utilize a big data run-time such as Tez, or Spark to execute SQL queries. This category includes systems such as Hive and Spark SQL.}
\item{ {\bf Native systems with MPP architectures:} These are systems that employ an MPP database architecture that use stand-alone long running processing to execute SQL queries over HDFS data. This category includes systems such as Impala, Presto, Drill, and Big SQL}
\item{{\bf Database hybrids:} These are systems that combine a relational SQL engine with Hadoop run-time. This category includes systems such as Hadapt, HAWQ, and Vortex.}
\item{ {\bf Database connectors:} This category includes the extensions provided by major database vendors to reach out to Hadoop and ingest HDFS data. }
\item{ {\bf noSQL/Hbase specific variants:} These are systems that provide a declarative SQL layer on top of noSQL systems, HBase in particular. The systems in this category are Phoenix, and Splice Machine. Note that these systems focus on OLTP-like workloads, whereas all other categories address OLAP-like analytics}
\end{itemize}

@@

Apache Drill \cite{drill} is an open-source project which aims at providing SQL-like declarative processing over self-describing semi-structured data. Its focus is on analyzing data without emposing a fixed schema or creating tables in a catalog like Hive MetaStore. It runs queries over files and HBase tables, and discovers data when reading input data. For every fixed chunk of data, it discovers its schema, creates an in-memory columnar representation, and generates specific code for processing. As such, it can accommodate data chunks with varying schemas. Drill provides a columnar execution engine that delays materialization as late as possible.
